\Title

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

%One advantage of our model is the generality of the possible kernels. However, this can be a limit as this increases the number of free parameters in the learning algorithm. As we observe that the timing of events in natural scene may be slightly variable (in the order of the synaptic time constant of $5~\ms$) an extension of the model would be to include a fixed filtering stage similar to that in the tempotron model~\citep{gutig2006tempotron}. This could be efficiently added as a temporal convolution on $\mathcal{C}^\postsynaddr(t)$.
%\noteAG{C'est pas clair pour moi ce que veut dire ou ce qu'ajoute le paragraphe précédent. La généralité des kernels est données dans le suivant et celui là fait appel à des concepts de filtrage pas très explicités. Aussi la référence au timing des events des scènes naturelles peut être justifié par une référence. }

We identified several limits to our model, the major of which we will now detail. First, the full framework is based on discrete binning of time which is incompatible with the analogous nature of biological time. We have used this binning to ease the implementation of the computational framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the time bin size and have shown it does not impact the results presented in this paper. This is consistent with the relative robustness of other  event-based frameworks such as HOTS (Lagorce 2017), where accuracy was not changed when jittering input spikes by up to 1ms of noise (grimaldi 2022). This hints at including analytically a precision term in the temporal value of incoming spikes, a mechanism potentially implemented by the approx 5 ms smoothing which is implemented at the level of the PSC of synapses. Furthermore, it is possible to overcome the problem using a purely even-driven based scheme. Indeed, it is not necessary to compute voltage males between two spikes (Morrison) and to define a purely event-driven framework. Such an architecture should provide with promising speed-ups for the computations.

A further limit is that the model that we defined is purely feed-forward. As a consequence, the spikes generated by post-synaptic neurons are produced solely based on the information in the classical receptive field. However, it is known that neurons in a layer may interact using lateral interaction (chavane 2022) and that this may be the basis of computational principles, for instance of binding of neighboring orientations which may be a basis for image categorization (Bednar). Moreover, neural information is modulated by feedback information, for instance do distinguish figure from background (Roelfsema 2016) and has proven to be essential in building realistic models of the primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022). It is not currently possible to implement these recurrent connections in our implementation (lateral or feedback), mainly because of our use of convolutions. However the generic theoretical model is capable of including these by inserting spikes in the list of spikes reaching the pre-synaptic addresses. While possible in theory, it should be properly adapted in practice such that these recurrent connections do not amplify neural activity outside an homeostatic state (by a death or a blow-up of activity).

Such recurrent activity would be essential to implement predictive or anticipative processes . This is essential in a neural system as it contains multiple different delays which necessitates a temporal alignment (Hoogendoorn 2019]. This was previously modeled to explain for instance the flash - lag visual illusion (Khoei 2017) As was previously started " ... " (Perrinet 2014], and our hetero-synaptic model would provide with a simple and elegant solution to that problem.

>>>>

%One advantage of our model is the generality of the possible kernels. However, this can be a limit as this increases the number of free parameters in the learning algorithm. As we observe that the timing of events in natural scene may be slightly variable (in the order of the synaptic time constant of $5~\ms$) an extension of the model would be to include a fixed filtering stage similar to that in the tempotron model~\citep{gutig2006tempotron}. This could be efficiently added as a temporal convolution on $\mathcal{C}^\postsynaddr(t)$.
%\noteAG{C'est pas clair pour moi ce que veut dire ou ce qu'ajoute le paragraphe précédent. La généralité des kernels est données dans le suivant et celui là fait appel à des concepts de filtrage pas très explicités. Aussi la référence au timing des events des scènes naturelles peut être justifié par une référence. }
We have identified several limits to our model, the most important being now detailed. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We tested the effect of the size of the time bin and showed that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS (Lagorce 2017), where the accuracy was not affected when the input spikes were subjected to noise up to 1ms (grimaldi 2022). This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the approximately 5 ms filtering that is implemented at the PSC of synapses. Furthermore, it is possible to overcome the problem of time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes ( Hanuschkin 2010) and thus define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations.

Another restriction is that this model is purely feed-forward. Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field. However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 (Chavane 2022) and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism (Bednar). Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background (Roelfsema 2016) and has been shown to be essential for building realistic models of primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022). It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions. However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses. While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a death or explosion of activity).

Such recurrent activity would be essential for implementing predictive or anticipatory processes. This is essential in a neural system as it contains multiple different delays that require temporal alignment (Hoogendoorn 2019). This has been previously modeled to explain for example the flash-lag illusion (Khoei 2017). As previously stated ".... " (Perrinet 2014), and our hetero-synaptic model would provide a simple and elegant solution to this problem.
